Follow-up Samples

### 2017-01-25 submission/manuscript_outline_20161024.Rmd
correct error with BH calculation

source('../code/functions.R')
source('../code/graphFunctions.R')

@@

***Previously Associated Cancer Bacteria***
We next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  First, the carcinoma samples showed a significant difference between initial and follow up samples for *Peptostreptococcus stomatis* (P-value = `r format(main_crc_bugs_summary["pept", "crc_BH"], digits = 3)`) and *Porphyromonas asaccharolytica* (P-value = `r format(main_crc_bugs_summary["porp", "crc_BH"], digits = 3)`) whereas there were no signficant differences in any of these OTUs in the adenoma samples [Table S4]. Second, when these OTUs were present, there was a clear magnitude difference based on whether they were from adenoma or carcinoma [Figure 2].  However, only a small percentage of those with adenoma or carcinoma were positive for any of these OTUs.

>>>>

source('../code/functions.R')

@@

***Previously Associated Cancer Bacteria***
We next examined whether there were differences in previously well described carcinoma associated OTUs.  These included the OTUs that aligned with *Porphyromonas asaccharolytica* (`r unique(filter(crc_bug_info, tax_id == "Porphyromonas")[, "otu"])`), *Fusobacterium nucleatum* (`r unique(filter(crc_bug_info, tax_id == "Fusobacterium")[, "otu"])`), *Parvimonas micra* (`r unique(filter(crc_bug_info, tax_id == "Parvimonas")[, "otu"])`), and *Peptostreptococcus stomatis* (`r unique(filter(crc_bug_info, tax_id == "Peptostreptococcus")[, "otu"])`).  First, the carcinoma samples showed a significant difference between initial and follow up samples for *Peptostreptococcus stomatis* (P-value = `r format(main_crc_bugs_summary["pept", "crc_BH"], digits = 3)`), *Parvimonas micra* (P-value = `r format(main_crc_bugs_summary["parv", "crc_BH"], digits = 3)`), and *Porphyromonas asaccharolytica* (P-value = `r format(main_crc_bugs_summary["porp", "crc_BH"], digits = 3)`) whereas there were no signficant differences in any of these OTUs in the adenoma samples. Second, when these OTUs were present, there was a clear magnitude difference based on whether they were from adenoma or carcinoma [Figure 2].  However, only a small percentage of those with adenoma or carcinoma were positive for any of these OTUs.

### 2017-01-25 submission/manuscript_outline_20161024.tex
correct error with BH calculation

\textbf{\emph{Previously Associated Cancer Bacteria}} We next examined
whether there were differences in previously well described carcinoma
associated OTUs. These included the OTUs that aligned with
\emph{Porphyromonas asaccharolytica} (Otu000153), \emph{Fusobacterium
nucleatum} (Otu000226), \emph{Parvimonas micra} (Otu000460), and
\emph{Peptostreptococcus stomatis} (Otu000653). First, the carcinoma
samples showed a significant difference between initial and follow up
samples for \emph{Peptostreptococcus stomatis} (P-value = 0.0183) and
\emph{Porphyromonas asaccharolytica} (P-value = 0.0154) whereas there
were no signficant differences in any of these OTUs in the adenoma
samples {[}Table S4{]}. Second, when these OTUs were present, there was
a clear magnitude difference based on whether they were from adenoma or
carcinoma {[}Figure 2{]}. However, only a small percentage of those with
adenoma or carcinoma were positive for any of these OTUs.

@@

Interestingly, when only previously associated cancer bacteria were
investigated only 2 out of the 4 had a significant decrease in relative
abundance between initial and follow up for the carcinoma group and 0
out of the 4 were significant for the adenoma group. This data suggests
that the changes that may be important in the bacterial microbiome after
surgery are not necessarily any one specific bacterium, a clear
depression of CRC related bacteria, or any addition of new or depressed
bacterium. Instead it may be due to a shift of the bacterial community
as a whole between existing members of the host's microbiome.

@@

\textbf{Figure 2: Previously Associated CRC Bacteria in Initial and
Follow up Samples.} A) Carcinoma initial and follow up samples. There
was a significant difference in initial and follow up sample for the
OTUs classfied as \emph{Peptostreptococcus stomatis} (P-value = 0.0183)
and \emph{Porphyromonas asaccharolytica} (P-value = 0.0154). B) Adenoma
initial and follow up samples. There were no significant differences
between initial and follow up.

>>>>

\textbf{\emph{Previously Associated Cancer Bacteria}} We next examined
whether there were differences in previously well described carcinoma
associated OTUs. These included the OTUs that aligned with
\emph{Porphyromonas asaccharolytica} (Otu000153), \emph{Fusobacterium
nucleatum} (Otu000226), \emph{Parvimonas micra} (Otu000460), and
\emph{Peptostreptococcus stomatis} (Otu000653). First, the carcinoma
samples showed a significant difference between initial and follow up
samples for \emph{Peptostreptococcus stomatis} (P-value = 0.00913),
\emph{Parvimonas micra} (P-value = 0.0185), and \emph{Porphyromonas
asaccharolytica} (P-value = 0.00913) whereas there were no signficant
differences in any of these OTUs in the adenoma samples. Second, when
these OTUs were present, there was a clear magnitude difference based on
whether they were from adenoma or carcinoma {[}Figure 2{]}. However,
only a small percentage of those with adenoma or carcinoma were positive
for any of these OTUs.

@@

Interestingly, when only previously associated cancer bacteria were
investigated only 3 out of the 4 had a significant decrease in relative
abundance between initial and follow up for the carcinoma group and 0
out of the 4 were significant for the adenoma group. This data suggests
that the changes that may be important in the bacterial microbiome after
surgery are not necessarily any one specific bacterium, a clear
depression of CRC related bacteria, or any addition of new or depressed
bacterium. Instead it may be due to a shift of the bacterial community
as a whole between existing members of the host's microbiome.

@@

\textbf{Figure 2: Previously Associated CRC Bacteria in Initial and
Follow up Samples.} A) Carcinoma initial and follow up samples. There
was a significant difference in initial and follow up sample for the
OTUs classfied as \emph{Peptostreptococcus stomatis} (P-value = 0.00913)
and \emph{Porphyromonas asaccharolytica} (P-value = 0.00913). B) Adenoma
initial and follow up samples. There were no significant differences
between initial and follow up.
